---
title: "Verification of the Centra Limit Theorem"
author: "rmmoya"
date: "October 26, 2019"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

nsim = 1000
nexp = 40
lambda = 0.02
```

## Overview

## Simulations


The exponential distribution is denoted as f(x) =  lambda * exp(-lambda*x)) and has the property that mean = 1/lambda and variance = 1/lambda^2.

We are going to perform **`r nsim`** simulations of **`r nexp`** exponentials and a **lambda=`r lambda`** and compare the resulting value of the estimated mean with **mean = 1/lambda=`r 1/lambda`**.

In order to use the same distribution of generated random numbers, a matrix of `r nsim`x`r nexp`, where each column contains the `r nexp` tests of the rexp function.

```{r simulations}
tests = matrix(data = NA,ncol = nsim, nrow = nexp)
for (i in (1:nsim)) tests[,i] = rexp(nexp, lambda) 
```


## Sample Mean versus Theoretical Mean

```{r sample_mean}

# Calculate the average mean from the simulations
mns = NULL
for (i in 1 : nsim) mns = c(mns, mean(tests[,i]))

# Real exponential distribution
# f(x) =  lambda * exp(-lambda*x))
exp_mean = 1/lambda

hist(mns)
abline(v = c(mean(mns), exp_mean), col = c('blue', 'red'), lty = c(1,2))
legend(x = "topright", c(paste0('Sample mean = ', round(mean(mns),2)), paste0('Theor. mean (1/lambda) = ', round(exp_mean,2))), lty = c(1,2), col = c('blue', 'red'))

```

The sample mean value obtained from the simulations is quite close to the theoretical mean, with **a difference of `r round(mean(mns) - exp_mean,2)`**, confirming the Central Limit Theorem.


## Sample Variance versus Theretical Variance
Repeating the exercise for the variance, we obtain the following distribution of sample variances:

```{r sample_variance}

# Calculate the average mean from the simulations
sigmas = NULL
for (i in 1 : nsim) sigmas = c(sigmas, var(tests[,i]))

# Real exponential distribution
# f(x) =  lambda * exp(-lambda*x))
exp_var = 1/lambda^2

hist(sigmas)
abline(v = c(mean(sigmas), exp_var), col = c('blue', 'red'), lty = c(1,2))
legend(x = "topright", c(paste0('Sample variance = ', round(mean(sigmas),2)), paste0('Theor.variance (1/lambda) = ', round(exp_var,2))), lty = c(1,2), col = c('blue', 'red'))

```

The sample variance value obtained from the simulations is quite close to the theoretical variance, with **a difference of `r round(mean(sigmas) - exp_var,2)`**, confirming the Central Limit Theorem.

## Distribution

The Central Limit Theorem states that the distribution of averages of iid random variables (properly normalized) becomes that of a standard normal as the sample size increases. 

In the following figure, it can be seen that the distribution of means from the simulations follows the normal distribution depicted with the solid line.

```{r}

hist(mns, freq = FALSE, ylim = c(0,0.06), main = "Distribution of sample means and the normal distribution", xlab = 'Sample means')
curve(dnorm(x, mean = 1/lambda, sd = 1/lambda/sqrt(nexp)), add = TRUE, lwd = 2, col = 'blue')
```

